Behavior Driven Development in Haskell

### 2012-05-25 paper.tex
Add subsubsections to Implementation -> doctest
More wording changes

This also means that typical usage examples is not provided by test code.

We agree that type signatures are good abstract documentation~\cite{free}, but
there are many functions whose behavior cannot be guessed from their names and
type signatures. A good example is \verb|nub :: [a] -> [a]| from {\tt Data.List}.

@@

In this paper, we explore how to apply BDD to Haskell.
We brought a test method from other programming language community and
integrated it with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck} as a BDD tool.
We also make use of \hspec{} inspired by rspec~\cite{rspec} for
BDD in Haskell.

The contribution of this paper is as follows:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problems}

In Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
and the {\tt test-framework} package integrates them.
We can categorize the {\tt test-framework} package into
a tool of TDD.
Before our work and \hspec{}, there is no BDD tool commonly used in Haskell community.

BDD tools used in other programming language communities are
mainly based on example test
and do not support property test.
To our experience, it is not difficult to find properties for
pure Haskell functions.
So, when we bring BDD tools to
Haskell community, we should integrate them with properties test.

We can categorize examples/properties into two types: for users and
and for developers.
Examples/properties for users are those which are useful as user documentation.
By contrast, examples/properties for developers are
those which are not useful as user documentation.
A good example of examples/properties for developers
is a corner example registered in a bug track system.

@@

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines markup syntax in Haskell's comment.
Haddock provides the code block markup indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block markup:

@@

\noindent Though the code block itslef is a necessary feature,
we think this kind usage has problems.
If the equations above are examples, we cannot distinguish
their results from their expression to be evaluated.
If they are properties, they should be distinguished
from example code.

@@

Here are summary of the problems:

\begin{itemize}
\item Haskell community does not have common syntax to describe
examples/properties with Haddock markup syntax.
\item Haskell community does not have a BDD tool for 
examples/properties for users.
\end{itemize}

@@

To specify examples/properties for users, we propose to add two new markup to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe examples/properties for developers, we propose to use \hspec{} also integrated with QuickCheck.

@@

@@

>>>>

This also means that typical usage examples are not provided by test code.

We agree that type signatures are good abstract documentation~\cite{free}, but
there are many functions whose behavior cannot be guessed from their names and
type signatures.  One example is \verb|nub :: [a] -> [a]| from {\tt Data.List}.

@@

In this paper, we explore how to apply BDD to Haskell.
We brought a test method from other programming language community and
integrated it with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
We also make use of \hspec{} inspired by RSpec~\cite{rspec} for
BDD in Haskell.

The contributions of this paper are:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem}

In the Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
and the {\tt test-framework} package integrates them.
We can categorize {\tt test-framework} as
a tool for TDD.
Before our work and \hspec{}, there was no widely used BDD tool for Haskell.

BDD tools used in other programming language communities are
mainly based on example tests,
and do not support property tests.
To our experience is, that it is not difficult to find properties for
pure Haskell functions.
So, when we bring BDD tools to
Haskell community, we should integrate them with properties test.

We can categorize examples/properties into two types: for users and
and for developers.
Examples/properties for users are those which are useful as user documentation.
By contrast, examples/properties for developers are
those which are not useful as user documentation.
A good example of examples/properties for developers
is a corner example registered in a bug tracking system.

@@

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines markup syntax in Haskell's comment.
Haddock provides markup for code blocks, indicated with ({\tt >}).
For example, the {\tt containers} packages has the following code block markup:

@@

\noindent Though the code block itself is a useful feature,
we think this kind of usage has problems.
If the equations above are examples, we cannot distinguish
their results from their expression to be evaluated.
If they are properties, they should be distinguished
from example code.

@@

Here is a summary of the problem:

\begin{itemize}
\item The Haskell community does not have common syntax to describe
examples/properties with Haddock markup syntax.
\item The Haskell community does not have a BDD tool for examples/properties
    for users.
\end{itemize}

@@

To be able to specify examples and properties for users, we propose to add new
markup to Haddock and to use {\tt doctest} integrated with QuickCheck. To
describe examples/properties for developers, we propose to use \hspec{} also
integrated with QuickCheck.

@@

\subsubsection{Testing examples}

@@

\subsubsection{Testing properties}
